---
title: Standardized dose-based weighting approaches for a drug-prescription network
keyword:
  - Polypharmacy
  - Network Analysis
  - Drug-Prescription Network
  - Dose-adjusted Weight
  - Prescription Registry

format:
  pdf:
    geometry:
      - top=2cm
      - bottom=2cm
      - left=2cm
      - right=2cm
    include-in-header:
      file: _header.tex

metadata-files:
  - _author.yml
  - _knitr-bib.yml
---

```{r init}

pkgs <- c("magrittr", "targets", "tarchetypes", "crew")
pkgs_load <- sapply(pkgs, library, character.only = TRUE)

options(digits = 2)
options(scipen = 999)

```

# Introduction

Situation:

- Drug-prescription network (DPN) is a method to evaluate multiple dyadic concurrence of prescribed medications
- Network science approach is applicable to a DPN
- Graph metrics can be used to describe the node, which represents single medication
- Centrality measures may attribute the relative importance of one medication compared to the others
- Graphical model can be used to evaluate the internal characteristic of a static and dynamic graph
- DPN is useful to evaluate polypharmacy

Problem:

- Polypharmacy regimes in psychiatric cases often combine an optimal dose medication with a sub-clinical dose medication
- The current approach of DPN only consider DDD = 1
- This will lead to loss of information if medications with DDD != 1 are not included
- We need a proper weighting method to incorporate dose into a DPN

Question:

- What is the suitable approach for weighting a DPN using DDD?

Resolution:

- Implement six weighting approaches
- Evaluate indirect consistency using ICC
- Evaluate direct consistency using TERGM

# Methods

- Source of data: IADB
- Apply weights
- Generate a DPN
- Measure graph metrics

## Weighting criteria

Previous studies on DPN always include medications with DDD of 1, which naturally results in an edge of 1 for each co-prescription in a patient. As such, a DPN has a baseline edges weight of $\omega_B$ that directly reflect the number of people having a particular co-prescription. The purpose of weighting a DPN is to adjust the edge weight using a standardized measure, e.g. with a defined daily dose (DDD). Since the weight is adjusted on an individual level, all baseline edge weight is set to $\omega_B$ = 1 for each approach. Six approaches used in this studies are resultant, product, quotient, absolute log, inverted absolute log, and density-based approach. Following criteria were followed when designing the weight adjustment approach:

1. The adjusted weight of co-prescribed medications with DDD of 1 is equal to the baseline weight
1. The adjusted weight has a range of $[0, \infty)$, limiting the weight to zero-inclusive definite real numbers

## Weighting approaches

The resultant-based approach is the simplest to calculate, where it generates the average of the dose and the baseline edge weight. The adjusted weight of $\omega_R$ is calculated by taking the resultant of DDD and the baseline weight $\omega_B$, then divide it by two, or more formally $\omega_R = (DDD + \omega_B) / 2$. Calculating the average produces a midpoint of the DDD and the baseline weight $\omega_B$. This approach has a proportionally linear relationship to the DDD, where medications with a higher DDD will have a higher weight.

The product-based approach is similar to the resultant-based approach, but it multiplies the dose to the baseline wedge weight. The adjusted weight of $\omega_P$ is calculated by taking the product of DDD and the baseline weight $\omega_B$, or more formally $\omega_P = DDD \cdot \omega_B$. Calculating the product produces DDD as a weight since $\omega_B$ = 1 on an individual-level data. The product-based approach is suitable in a scenario where DDD reflects the importance of a combination.

The quotient-based approach is an inversion of the product-based approach, where it divides the baseline wedge weight to the dose. The adjusted weight of $\omega_Q$ is calculated by taking the quotient of the baseline weight $\omega_B$ and DDD, or more formally $\omega_Q = \frac{\omega_B}{DDD}$. Calculating the quotient produces a weight that will highlight combinations with sub-dose medications. The quotient-based approach is intended for a scenario where a DPN is used to evaluate augmentation polypharmacy, that is the use of sub-dose medication and an optimal-dose medication to treat the same clinical condition [@Shrivastava2013].

The absolute log approach takes the absolute difference between the baseline weight $\omega_B$ and the absolute log of DDD, or more formally $\omega_L = |\omega_B - |\log_{10}(DDD)||$. The absolute log approach is considered because the value of $\log(1)$ is zero for all log bases. Therefore, a difference of the baseline weight and the log of DDD is $\omega_B$ when DDD = 1, fulfilling the first criteria. The absolute of $\log$ reverts the negative values when DDD < 1, and thus, keeping the weight $\omega_L$ less than 1. The absolute log approach is suitable when we have a hypothesized maximum of a DDD, which can be set as the log base. Values of DDD beyond the log base will increase the weight $\omega_L$ proportionally. In this study, the log base is conveniently set to $\log_{10}$ assuming that DDD = 10 as the theoretical maximum for any medication in a population.

The inverted log approach inverts absolute-log weight $\omega_L$, or more formally $\omega_I = |\omega_B - \omega_L| + \omega_B$. The main motivation of using an inverted log approach is to increase the weight given that DDD $\neq$ 1. Since it uses $\omega_L$ to calculate the weight, it shares the same limitation as well. For the absolute log approach with the log base is set to 10, DDD of 0.1 becomes the turning point, as shown in @fig-plt-sim. Since 0.1 is the lowest DDD in the dataset, we changed its value to 0.11 to avoid the turning-point problem. Despite its limitation, the inverted log approach is usable to increase the weight when DDD $\neq$ 1.

The density-based approach relies on the probability density function of a normal distribution, assuming that the weight follows a normal distribution, $\omega_D \sim \mathcal{N}(\mu, \sigma)$. The parameter mean $\mu$ is set to DDD and standard deviation $\sigma$ is set to $\frac{\omega_B}{3}$ to follow the empirical rule of 68-95-99.5 [@grafarend2006linear]. Since the density when DDD = 1 may exceed 1 and violates the first criteria, the weight $\omega_D$ is rescaled by dividing with the maximum value. Density-based approach symmetrically diminishes the weight when DDD drifts away from 1.

| Approach | Formula | Use case   |
| :------- | :------ | :--------- |
| Baseline | $\omega_B = 1$ | Set the baseline edge weight in a DPN, where the weight $\omega_B$ is set to 1 for each co-prescription in a patient |
| Resultant | $\omega_R = \frac{DDD + \omega_B}{2}$ | Proportionally adjust weight based on dose |
| Product   | $\omega_P = DDD \cdot \omega_B$ | Aggressively adjust weight based on dose; combination of high-dose medications will have the highest weight |
| Quotient  | $\omega_Q = \frac{DDD}{\omega_B}$ | Highlight combinations with sub-dose medications, e.g. to evaluate augmentation polypharmacy |
| Absolute log | $\omega_L = |\omega_B - | \log_{10}(DDD) ||$ | Put less weight for DDD other than 1 |
| Inverted $\omega_L$ | $\omega_I = |\omega_B - \omega_L| + \omega_B$ | Put more weight for DDD other than 1 |
| Density | $\omega_D = \frac{1}{\frac{\omega_B}{3} \sqrt{2 \pi}} e^{-\frac{1}{2}\left(\frac{DDD - \omega_B}{\omega_B/3}\right)^2}$ | Symmetrically assign diminished weight for DDD drifting away from 1 |

: Weighting approaches using defined daily dose (DDD) for a drug-prescription network {#tbl-approach tbl-colwidths="[15,30,55]"}

```{r}
#| label: fig-plt-sim
#| fig-cap: Simulation on all weighting approaches. The dashed line indicates DDD = $\omega$ = 1.
#| fig-height: 2
#| fig-width: 10

tar_read(plt_sim)

```

# Results

{{< include results/_summary.qmd >}}

{{< include results/_cor.qmd >}}

{{< include results/_icc.qmd >}}

# Discussion

# Conclusion

# Reference

